---
title: "STAT 5361 - Homework 3"
# subtitle: "possible subtitle goes here"
author:
  - Jin Xu, Jiayi Liu
date: "`r format(Sys.time(), '%d %B %Y')`"
documentclass: article
papersize: letter
fontsize: 12pt
bibliography: template.bib
biblio-style: datalab
keywords: Template, R Markdown, bookdown, Data Lab
# keywords set in YAML header here only go to the properties of the PDF output
# the keywords that appear in PDF output are set in latex/before_body.tex
output: 
  pdf_document:
    latex_engine: xelatex
    pandoc_args: [
      "--number-sections",
      "--number-offset=1"
    ]
---



```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
## some utility functions, see the source code for details
source("utils_template.R")

## specify the packages needed
pkgs <- c("splines2", "DT", "webshot", "leaflet", "graphics")
need.packages(pkgs)

## external data can be read in by regular functions,
## such as read.table or load

## get output format in case something needs extra effort
outFormat <- knitr::opts_knit$get("rmarkdown.pandoc.to")
## "latex" or "html"

## for latex and html output
isHtml <- identical(outFormat, "html")
isLatex <- identical(outFormat, "latex")
latex <- ifelse(isLatex, '\\LaTeX\\', 'LaTeX')

## specify global chunk options
knitr::opts_chunk$set(fig.width = 5, fig.height = 4, dpi = 300,
                      out.width = "90%", fig.align = "center")

```

### Problem 1 {-}
\begin{align}
    &n(y_i - x_i^T\beta_j^{(k)};0,\sigma^{2(k)})\\
    &= \dfrac{(2\pi)^{-d/2}}{\sqrt{|\sigma^{2k}|}}\exp\{(-\dfrac{1}{2}(y_i - x_i^T\beta_j^{(k)})(\sigma^2)_{k}^{-1}(y_i - x_i^T\beta_j^{(k)})\}\\
    &Q(\Psi | \Psi^{(k)}) \\    
    &= \sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1}\ln[(2\pi)^{-d/2}\pi_j^{(k)}] - \dfrac{1}{2}\sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1}\ln|\sigma^{2k}|- \dfrac{1}{2}\sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1}(y_i - x_i^T\beta_j^{(k)})(\sigma^2)_{k}^{-1}(y_i - x_i^T\beta_j^{(k)}) \\
    &= I_1 - \dfrac{I_2}{2} - \dfrac{I_3}{2} \\
    &I_1 = \sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1}\ln[(2\pi)^{-d/2}\pi_j^{(k)}] \\
    &I_2 = \sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1}\ln|\sigma^{2k}| \\
    &I_3 = \sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1}(y_i - x_i^T\beta_j^{(k)})(\sigma^2)_{k}^{-1}(y_i - x_i^T\beta_j^{(k)})\\
    &\beta_j^{(k+1)} = \dfrac{\sum_{i=1}^{n}{\rm x}_iy_ip_{ij}^{(k+1)}}{\sum_{i=1}^{n}{\rm x}_i{\rm x}_i^{(T)}p_{ij}^{(k+1)}}, j=1, ..., m\\
    &S_k = \sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1}\ln|\sigma^{2k}| + \sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1}(y_i - x_i^T\beta_j^{(k)})(\sigma^2)_{k}^{-1}(y_i - x_i^T\beta_j^{(k)})\\
    &\sigma^{2(k+1)} = \dfrac{\sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1}(y_i - {\rm x}_i^T\beta_j^{(k+1)})^2}{n}\\
    &I_1 = \sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1}\ln[(2\pi)^{-d/2}\pi_j^{(k)}]\\
    &=\sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1}\ln[(2\pi)^{-d/2}] + \sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1}\ln[\pi_j^{(k)}]\\
    &=-(d/2) \ln(2\pi)\sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1} + \sum_{i=1}^{n}\sum_{j=1}^{m}p_{ij}^{k+1}\ln[\pi_j^{(k)}]\\
    &\pi_j^{(k+1)}= \dfrac{\sum_{n}^{i=1}p_{ij}^{(k+1)}}{n}\\
\end{align}

### Problem 2 {-}

### Problem 2(a) {-}

We want $C \int_{0}^ \infty (2x^{\theta - 1}+x^{ \theta- \frac{1}{2} })*e^{-x}{ \rm d}x =1$ And we know that$\Gamma( \alpha)=\int_{0}^ \infty x^{\alpha - 1}*e^{-x}{ \rm d}x$. Therefore,
\begin{align}
    &C[ \int_{0}^ \infty 2x^{ \theta-1}*e^{-x}{ \rm d}x + \int_{0}^ \infty x^{ \theta- \frac{1}{2}}*e^{-x}{ \rm d}x]= C(2* \Gamma( \theta)+ \Gamma( \theta+ \frac{1}{2}))=1\\
    &C= \frac{1}{2 \Gamma( \theta)+ \Gamma( \theta +\frac{1}{2})}\\
    &g(x)= \frac{1}{2 \Gamma( \theta)+ \Gamma( \theta +\frac{1}{2})}(2x^{\theta - 1}*e^{-x}+x^{ \theta- \frac{1}{2} }*e^{-x})\\    
    &=\frac{1}{2 \Gamma( \theta)+ \Gamma( \theta +\frac{1}{2})}(2 \Gamma( \theta)*f(x; \theta,1)+ \Gamma( \theta+ \frac{1}{2})*f(x; \theta + \frac{1}{2},1))\\
    &= \frac{2 \Gamma( \theta)}{2 \Gamma( \theta)+ \Gamma( \theta +\frac{1}{2})}*f(x; \theta,1)+\frac{\Gamma( \theta+ \frac{1}{2})}{2 \Gamma( \theta)+ \Gamma( \theta +\frac{1}{2})}*f(x; \theta + \frac{1}{2},1)\\
\end{align}

f(x:$\theta$,1) is a gamma distribution with $\alpha=\theta;\beta=1$;

f(x:$\theta+ \frac{1}{2}$,1) is a gammadistribution with $\alpha=\theta+\frac{1}{2};\beta=1$.

So, $g(x)$ is a mixture of two Gamma distributions. The weights of these two distributions are shown in the fuction above.

### Problem 2(b) {-}

Frist, we set $\theta=3$. The weights for$f(x;\theta,1)$ and $f(x;\theta + \frac{1}{2},1)$ are $0.712155$ and $0.287845$ respectively.
```{r}
integrand1 <- function(x) {2*x^2*exp(-x)}
integrand2 <- function(x) {x^(5/2)*exp(-x)}
in1 <- (integrate(integrand1, lower = 0, upper = Inf))
in2 <- (integrate(integrand2, lower = 0, upper = Inf))
print(in1)
print(in2)
```
```{r}
N <- 10000
components <- sample(1:2, prob=c(0.712155,0.287845),size=N,replace=TRUE)
sps <- c(3, 7/2)
samples <- rgamma(n=N,shape = sps[components])
d1 <- density(samples)
plot(d1,col="darkblue", main = "kernel density estimation and true density")
curve((0.712155 * dgamma(x, 3) + 0.287845 * dgamma(x, 7/2)), col = "darkred", add = TRUE)
legend('topright', c('sample density','true density'), lwd = 2, col = c('darkblue','darkred'))
```

### Problem 2(c) {-}

```{r}

# set theta = 3

sample2 <- samples
envelope2 <- function(x){
  (0.712155 * dgamma(x, 3) + 0.287845 * dgamma(x, 7/2))
}

target2 <- function (x)
{
  (4 + x)^(1/2) * x^2* exp(-x)
}
accept = c()
rejection_sampling <- function(sample2, envelope2, target2){
  for(i in 1:length(sample2)){
    U = runif(1,0,1)
    if(envelope2(sample2[i])*U*1<= target2(sample2[i])) {
      accept[i] = 'Yes'}
    else {
      accept[i] = 'No'
    }
  }
}
T2 <- data.frame(sample2, accept = 'Yes')
hist(T2[,1][T2$accept=='Yes'], freq = FALSE, breaks = seq(0,20,0.5), main = 'estimated density of a random sample', xlab = 'X')
lines(density(sample2), col = "darkblue")

```


### Problem 3 {-}

### Problem 3(a) {-}

Let U~Unif(0,1), for any $\theta, \beta >0$, $U^{1/\theta} \sim \rm Beta(\theta,1)$, $1-U^{1/\beta} \sim \rm Beta(1, \beta)$

The density of the distributions are $g_\theta(x) = \theta x^{\theta-1}$, $g_\beta(x) = \beta (1-x)^{\beta-1}$

Therefore, if $g(x)$ is a mixture of $g_\theta(x)$ and $g_\beta(x)$, $g(x) = p_1g_\theta(x) +p_2g_\beta(x)$, 

$p_1 =\dfrac{1}{1+x^2}\rm B(\theta,1)$, 

$p_2 =\sqrt{2+x^2} \rm B(1,\beta)$.

Assume $\theta = 2$, $\beta=3$: 

```{r}
p1 <- beta(2,1)
p2 <- beta(1,3)*3^(1/2)
N3 <- 10000
components3 <- sample(1:2, prob=c(p1,p2),size=N3,replace=TRUE)
sps1 <- c(2, 1)
sps2 <- c(1, 3)
sample3 <- rbeta(n=N3,shape1=sps1[components3], shape2 = sps2[components3])
envelope3 <- function(x){
  dbeta(x,shape1=sps1[components3], shape2 = sps2[components3])
}
target3 <- function(x){
  x/(1+x^2)+(2+x^2)^(1/2)*(1-x)^2
}
accept = c()
rejection_sampling <- function(sample3, envelope3, target3){
  for(i in 1:length(sample3)){
    U = runif(1,0,1)
    if(envelope3(sample3[i])*U <= target3(sample3[i])) {
      accept[i] = 'Yes'}
    else {
      accept[i] = 'No'
    }
  }
}
T3 = data.frame(sample3, accept = 'Yes')
plot(density(sample3),main="Estimated density of a random sample",col = "darkblue")
```


### Problem 3(b) {-}

$p_1 =\dfrac{1}{1+x^2}\rm B(\theta,1)$, 

$p_2 =\sqrt{2+x^2} \rm B(1,\beta)$.

Assume $\theta = 2$, $\beta=3$: 

```{r}
p1 <- beta(2,1)
p2 <- beta(1,3)*3^(1/2)
p <- p1 + p2
N <- 10000

#Sample N random uniforms U
U <- runif(N)

#Variable to store the samples from the mixture distribution                                             
rand.samples <- rep(NA,N)

#Sampling from the mixture
for(i in 1:N){
    if (U[i] < p1/p) {
        rand.samples[i] <- rbeta(1,2,1)
    }else{
        rand.samples[i] <- rbeta(1,1,3)
    }
}
#rejection-sampling
rejection_sampling <- function(rand.samples, envelope3, target3){
  for(i in 1:length(rand.samples)){
    U = runif(1,0,1)
    if(envelope3(rand.samples[i])*U <= target3(rand.samples[i])) {
      accept[i] = 'Yes'}
    else {
      accept[i] = 'No'
    }
  }
}
T4 <- data.frame(rand.samples, accept = 'Yes')
#Density plot of the random samples

plot(density(rand.samples),main="Density Estimate of the Mixture Model",col = "darkblue")
```










